---
title: "HW6 - Exercise6"
author: JooChul Lee
date: "`r format(Sys.time(), '%d %B %Y')`"
documentclass: article
knit: "bookdown::render_book('Exercise2.Rmd', 'bookdown::pdf_document2')"
fontsize: 11pt
papersize: letter
header-includes:
   - \usepackage{algorithm}
   - \usepackage{algpseudocode}
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## for latex and html output
isHtml <- knitr::is_html_output()
isLatex <- knitr::is_latex_output()
latex <- ifelse(isLatex, '\\LaTeX\\', 'LaTeX')

## specify global chunk options
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
                      out.width = "90%", fig.align = "center")

```

# Rejection sampling

## Find the value of the normalizing constant. Show that $g$ is a mixture of Gamma distributions. Identify the component distributions and their weights in the mixture.
  \begin{align}
    C\int_0^\infty (2 x^{\theta-1} +  x^{\theta-1/2}) e^{-x} dx 
    &= C (\int_0^\infty 2 x^{\theta-1} e^{-x} dx + \int_0^\infty x^{\theta-1/2} e^{-x} dx)\\
    &= C (2 \Gamma(\theta)  + \Gamma(\theta + \frac{1}{2})) \\
    &=1
  \end{align}
 
Thus, $C = \dfrac{1}{2 \Gamma(\theta)  + \Gamma(\theta + \frac{1}{2})}$

Then, we know that
$$ g(x) =\dfrac{2\Gamma(\theta)}{2 \Gamma(\theta)  + \Gamma(\theta + \frac{1}{2})} Gamma(\theta,1) + \dfrac{\Gamma(\theta + \frac{1}{2})}{2 \Gamma(\theta)  + \Gamma(\theta + \frac{1}{2})} Gamma(\theta + \frac{1}{2},1)$$
which is the mixture gamma distribution. The component distributions are $Gamma(\theta,1)$ and $Gamma(\theta + \frac{1}{2},1)$. The weights are $\dfrac{2\Gamma(\theta)}{2 \Gamma(\theta)  + \Gamma(\theta + \frac{1}{2})}$ and $\dfrac{\Gamma(\theta + \frac{1}{2})}{2 \Gamma(\theta)  + \Gamma(\theta + \frac{1}{2})}$.
\newpage

## Design a procedure (pseudo-code) to sample from $g$; implement it in an R function

\begin{algorithm}
\caption{procedure to sample from $g$}
\begin{algorithmic}[1]
\Function{}{n, $\theta$}
\Comment{n : sample size, $\theta$ : parameter}
\State u : Draw samples with size n from standard uniform distribution
\For{i in 1:n} \Comment{Loop by data size}
\If{$u < \dfrac{2\Gamma(\theta)}{2 \Gamma(\theta)  + \Gamma(\theta + \frac{1}{2})}$}
\State $x_i$ : Draw a sample from $Gamma(\theta,1)$
\Else
\State $x_i$ : Draw a sample from $Gamma(\theta + \frac{1}{2},1)$
\EndIf
\EndFor
   \State \textbf{return} $x$ : sample with size n 
\EndFunction
\end{algorithmic}
\end{algorithm}

```{r}
sampling = function(n, theta)
{
   u = runif(n)
   Num = sum( u < 2*gamma(theta)/( 2*gamma(theta) + gamma(theta + 0.5) ) )
   sample = c(rgamma(Num, theta, 1), rgamma(n-Num, theta+0.5, 1))
   return(sample)
}
theta = 1
integrand = function(x)
{
   sqrt(4+x)*x^(theta-1)*exp(-x)
}
integral = integrate(integrand, 0, Inf)
aa = sampling(10000, theta)
hist(aa,freq=F,ylim = c(0,0.7),xlab='',main='')
lines(density(aa),col="red",lwd=4)
curve(sqrt(4+x)*x^(theta-1)*exp(-x)/as.numeric(integral[1]), add = T,col="blue",lwd=2)
legend('topright', legend=c("Simulated","Actual"),lty=c(1,1),
       lwd=c(2.5,2.5),col=c('red','blue'))
```
The figure is the result of $\theta=1$ and n=10000.

## Design a procedure (pseudo-code) to use rejection sampling to sample from $f$ using $g$ as the instrumental distribution.

First of all, we need to choose $\alpha$ such that 
$$q(x) \le \alpha g(x)$$
where $q(x) = \sqrt{4+x}x^{\theta -1}e^{-x}$

Since $\sqrt{4+x} \le \sqrt{4} + \sqrt{x}$,
$$q(x) = \sqrt{4+x}x^{\theta -1}e^{-x} \le (2 x^{\theta-1} +  x^{\theta-1/2}) e^{-x}= \frac{1}{C} g(x)$$
Then, $\alpha$ could be $\frac{1}{C} = 2 \Gamma(\theta)  + \Gamma(\theta + \frac{1}{2})$

Thus, we can choose $\alpha = \frac{1}{C}$ which satisfies the condition $q(x) \le \alpha g(x)$.

\begin{algorithm}
\caption{procedure to sample from $f$}
\begin{algorithmic}[1]
\Function{}{n, $\theta$}
\Comment{n : sample size, $\theta$ : parameter}
\State u : Draw one sample from standard uniform distribution
\State x : Draw one sample from mixture dist. based on $g(x)$
\If{$u > \dfrac{q(x)}{\alpha g(x)}$}
\State Go step 2-3
\Else
\State return x
\EndIf
\State Repeat the steps until x's size is n
   \State \textbf{return} $x$ : sample with size n 
\EndFunction
\end{algorithmic}
\end{algorithm}

```{r}
sampling = function(n, theta)
{
   u = runif(n)
   Num = sum( u < 2*gamma(theta)/( 2*gamma(theta) + gamma(theta + 0.5) ) )
   sample = c(rgamma(Num, theta, 1), rgamma(n-Num, theta+0.5, 1))
   return(sample)
}
algorithm2 = function(n, theta)
{
   sample = c()
   while( length(sample) <= n )
   {
      u = runif(1); x = sampling(1, theta)
      if(u <= sqrt( x + 4)/(sqrt(x)))
         sample= c(sample, x)
   }
   return(sample)
}
theta = 1
integrand = function(x)
{
   sqrt(4+x)*x^(theta-1)*exp(-x)
}
integral = integrate(integrand, 0, Inf)
aa = algorithm2(10000, theta)
hist(aa,freq=F,ylim = c(0,0.7),xlab='',main='')
lines(density(aa),col="red",lwd=4)
curve(sqrt(4+x)*x^(theta-1)*exp(-x)/as.numeric(integral[1]),add = T,col="blue",lwd=2)
legend('topright', legend=c("Simulated","Actual"),lty=c(1,1),
       lwd=c(2.5,2.5),col=c('red','blue'))
```
The figure is the result of $\theta=1$ and n=10000.

# Mixture Proposal

## Design a procedure (pseudo-code) to sample from $f$ using a mixture of Beta distributions as the instrumental density.

First of all, we know that

$$\frac{x^{\theta-1}}{1+x^2} + \sqrt{2+x^2}(1-x)^{\beta-1} \le x^{\theta-1} + 2(1-x)^{\beta-1}$$
Denote that $q(x) = \frac{x^{\theta-1}}{1+x^2} + \sqrt{2+x^2}(1-x)^{\beta-1}$ and $g(x) \propto x^{\theta-1} + 2(1-x)^{\beta-1}$

Then, 

$g(x) = \frac{\bold{B}(\theta,1)}{\bold{B}(\theta,1)+2\bold{B}(1,\beta)} Beta(\theta,1) + \frac{2\bold{B}(1,\beta)}{\bold{B}(\theta,1)+2\bold{B}(1,\beta)}Beta(1,\beta)$
Thus, we get that
$$p_1 = \frac{\bold{B}(\theta,1)}{\bold{B}(\theta,1)+2\bold{B}(1,\beta)}$$
$$p_2 = \frac{2\bold{B}(1,\beta)}{\bold{B}(\theta,1)+2\bold{B}(1,\beta)}$$

$$g_1 = Beta(\theta,1)$$
$$g_2 = Beta(1,\beta)$$
Let $\alpha = \bold{B}(\theta,1)+2\bold{B}(1,\beta)$.

Then, $$q(x) \le \alpha g(x)$$

\begin{algorithm}
\caption{procedure to sample from $f$}
\begin{algorithmic}[1]
\Function{}{n, $\theta$}
\Comment{n : sample size, $\theta$ : parameter}
\State u : Draw one sample from standard uniform distribution
\State x : Draw one sample from mixture dist. based on $g(x)$
\If{$u > \dfrac{q(x)}{\alpha g(x)}$}
\State Go step 2-3
\Else
\State return x
\EndIf
\State Repeat the steps until x's size is n
   \State \textbf{return} $x$ : sample with size n 
\EndFunction
\end{algorithmic}
\end{algorithm}

```{r}
sampling1 = function(n, theta, beta)
{
   u = runif(n)
   Num = sum( u < beta(theta, 1)/( 2*beta(1, beta) + beta(theta, 1) ) )
   sample = c(rbeta(Num, theta, 1), rbeta(n-Num, 1, beta))
   return(sample)
}
algorithm3 = function(n, theta, beta)
{
   sample = c()
   while( length(sample) <= n )
   {
      u = runif(1); x = sampling1(1, theta, beta)
      q = (x^(theta - 1)/(1+x^2)) + sqrt(2+x^2)*(1-x)^(beta-1)
      g = x^(theta -1) + 2*(1-x)^(beta-1)
      if(u <= q/g)
         sample= c(sample, x)
   }
   return(sample)
}
theta = 5; beta = 10
integrand = function(x)
{
   (x^(theta - 1)/(1+x^2)) + sqrt(2+x^2)*(1-x)^(beta-1)
}
integral = integrate(integrand, 0, 1)
aa = algorithm3(10000, theta, beta)
hist(aa,freq=F,xlab='',main='')
lines(density(aa),col="red",lwd=4)
curve( ((x^(theta - 1)/(1+x^2)) + sqrt(2+x^2)*(1-x)^(beta-1))/as.numeric(integral[1]),
       add = T,col="blue",lwd=2)
legend('topright', legend=c("Simulated","Actual"),lty=c(1,1),
       lwd=c(2.5,2.5),col=c('red','blue'))
```
The figure is the result of $(\theta=5, \beta=10)$ and n=10000.

## Design a procedure (pseudo-code) to do this; implement it with an R function.

$$\frac{x^{\theta-1}}{1+x^2}, \quad \sqrt{2+x^2}(1-x)^{\beta-1}$$

We can see that
$$\frac{x^{\theta-1}}{1+x^2} \le x^{\theta-1}$$
$$\sqrt{2+x^2}(1-x)^{\beta-1} \le 2(1-x)^{\beta-1}$$

Then,
$$q_1 = \frac{x^{\theta-1}}{1+x^2},\; g_1 =Beta(\theta,1),\; \alpha_1 = \bold{B}(\theta,1) $$
$$q_2 = \sqrt{2+x^2}(1-x)^{\beta-1},\; g_2 = Beta(1,\beta),\; \alpha_2 = 2\bold{B}(1,\beta) $$

\begin{algorithm}
\caption{procedure to sample from $f$}
\begin{algorithmic}[1]
\Function{}{n, $\theta$}
\Comment{n : sample size, $\theta$ : parameter}
\State Sample k from \{1,2\}
\State u : Draw one sample from standard uniform distribution
\State x : Draw one sample from dist. based on $g(x)_k$
\If{$u > \dfrac{q_k(x)}{\alpha_k g_k(x)}$}
\State Go step 2-3
\Else
\State return x
\EndIf
\State Repeat the steps until x's size is n
   \State \textbf{return} $x$ : sample with size n 
\EndFunction
\end{algorithmic}
\end{algorithm}

```{r}
algorithm4 = function(n, theta, beta)
{
   sample = c()
   while( length(sample) <= n )
   {
      k =sample(1:2,1)
      u = runif(1); 
      if(k == 1)
      {   
         x = rbeta(1,theta, 1)
         q = (x^(theta - 1)/(1+x^2))
         g = x^(theta -1) 
         if(u <= q/g)
            sample= c(sample, x)
      }   
      else {
         x = rbeta(1, 1, beta)
         q = sqrt(2+x^2)*(1-x)^(beta-1)
         g = 2*(1-x)^(beta-1)
         if(u <= q/g)
            sample= c(sample, x)
         }
   }
   return(sample)
}
theta = 5; beta = 10
integrand = function(x)
{
   (x^(theta - 1)/(1+x^2)) + sqrt(2+x^2)*(1-x)^(beta-1)
}
integral = integrate(integrand, 0, 1)
aa = algorithm4(10000, theta, beta)
hist(aa,freq=F,xlab='',main='')
lines(density(aa),col="red",lwd=4)
curve( ((x^(theta - 1)/(1+x^2)) + sqrt(2+x^2)*(1-x)^(beta-1))/as.numeric(integral[1]),
       add = T,col="blue",lwd=2)
legend('topright', legend=c("Simulated","Actual"),lty=c(1,1),
       lwd=c(2.5,2.5),col=c('red','blue'))
```
The figure is the result of $(\theta=5, \beta=10)$ and n=10000.